On Sample Size for Block Bootstrap Confidence Intervals to Have Desired Coverage Rates

### 2023-08-21 Manuscript/manuscript.tex
smoothed simulation and conclusion

\jy{Can we come up with a good name for the proposed method? Other people may cite this.}\mc{bias-centered? i'm not sure}
\paragraph{A Proposed CI}
We propose a CI that is centered at the point estimator and uses the variation
in the bootstrap estimates to construct the error bound. This interval requires
the computation of $\bar\theta_n^{(b)}$, the mean of all bootstrap point
estimates, and $\widehat{\text{bias}} = \bar\theta_n^{(b)} - \hat\theta_n$. A
$(1 - \alpha)100\%$ CI is
\[
(\hat\theta_{n, \alpha/2}^{(b)} - \widehat{\text{bias}}, \quad
\hat\theta_{n, 1 - \alpha/2}^{(b)} - \widehat{\text{bias}}).
\]
This CI is centered around $\hat\theta_n$ and can also be rewritten
as 
\[
(\hat\theta_n + \hat\theta_{n, \alpha/2}^{(b)} - \bar\theta_n^{(b)}, \quad
\hat\theta_n + \hat\theta_{n, 1 - \alpha/2}^{(b)} - \bar\theta_n^{(b)}).
\]
It is asymmetric, as different critical values are used to compute
the lower and upper bounds.

@@

We designed a simulation study to compare the performance of the different
block bootstrap CI methods. In particular, we generated time series $X_t$ from
a 1st order auto-regressive (AR(1)) process:
\[
X_t = \phi X_{t-1} + \epsilon_t,
\]
where $\phi$ is an auto-regressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: $\{-0.4, -0.2, 0.0, 0.2, 0.4\}$. The
series $X_t$ has mean zero and variance
$\sigma_x^2 = (1 - \phi^2) \sigma_{\epsilon}^2$, so for each value of $\phi$,
we set $\sigma_{\epsilon}^2 = 1 / (1 - \phi^2)$ such that $\sigma_x^2 = 1$.

@@

\jy{The results discussion need to be streamlined. The three paragraphs are too
  similar, and only student's ci is discussed. Questions to consider:
  1) which methods give the right coverage as sample size increases?
  2) which methods needs smaller sample sizes to get the right coverage?
  3) what's the difference between the results in positive and negative
  dependence?
  4) how does the result change as dependence gets stronger?
  } \mc{addressed}

For estimating the mean parameter~$\mu$, Figure~\ref{fig:mu} suggests that all
methods eventually approach correct coverage of $\mu$ as sample size
increases. Student's $t$ CIs generally does not need relatively large sample
sizes to achieve correct coverage, except for samples with strong negative
dependence, in which case, they actually over-cover $\mu$ for smaller sample
sizes. For instance, for a sample with $n = 100$ and $\phi = -0.4$, the lower
bound for a Students $t$ CIs coverage of $\mu$ is greater than 95\%, whereas
the coverage intervals for other methods contain 95\%. In fact, Standard,
Percentile, $BC$, and $BC_a$, and Proposed Method CIs require similar sample
sizes to to recover $\mu$ at the nominal level for all combinations of $n$ and
$\phi$. All methods seem to require a smaller sample to recover $\mu$ at the
nominal rate when dealing with negative dependence versus positive dependence.
For example, $BC$ CIs recovers $\mu$ for $n \geq 100$ when $\phi = -0.2$, but
they only recover $\mu$ for $n \geq 800$ when $\phi = 0.2$. In addition, as a
negative dependence gets stronger, holding everything else equal, coverage
increases, which lead to the Student $t$ CI's aforementioned over-coverage. As
a positive dependence gets stronger, holding everything else equal, coverage
decreases, and a larger sample is necessary to recover $\mu$.

For estimating the standard deviation parameter~$\sigma_x$,
Figure~\ref{fig:sigma} suggests that each method can reach nominal coverage of
$\sigma_x$ if the sample is large enough, but for a given $n$ and $\phi$,
coverage of $\sigma_x$ will be lower than coverage of $\mu$ in general. Like
$\mu$, $\sigma_x$ can be covered by Student $t$ CIs with smaller sample sizes
when compared to other methods, and there is no over-coverage issue for
$\phi = -0.4$. Standard, Percentile, $BC$, $BC_a$, and Proposed Method CIs
again have similar performance. All methods seem to have slightly higher
coverage of $\sigma_x$ when $\phi$ is negative versus when $\phi$ is positive.
Regardless of the sign, coverage of $\sigma_x$ gets worse as the strength of
the temporal dependence increases.

@@

To summarize, the CI type chosen should be based on the target parameter. When
estimating $\mu$ and $\sigma_x$, any CI will do, although Student's $t$ CIs
perform noticeably better than the others. However, when estimating $\phi$,
the choice of method is of utmost importance as to avoid coverage
deterioration. A smaller sample size is generally required to estimate a
parameter for a sample with a negative $\phi$ versus a positive $\phi$ of the
same magnitude. In order to know if coverage will increase as the strength of
the temporal dependence increases, one must know what the parameter of
interest is, and in the case of $\mu$: $\phi$'s sign.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/norm_phi_intervals}
  \caption{100 replicate Percentile, Bias-Corrected, and Proposed Method
    Bootstrap CIs for samples of $n = 800$ for an AR(1) process with
    $\phi = 0.4$. For each replicate, the lower and upper bounds of the CIs
    are displayed, as well as $\hat\theta_n$ (blue) and $\bar\theta_n^{(b)}$
    (red). }
  \label{fig:npi}
\end{figure}

These results beg the question of why certain methods perform poorly. To
address this question, plots of 100 CIs using each of the Percentile, $BC$,
and Proposed methods were constructed. As shown in Figure~\ref{fig:npi},
Percentile CIs appear to suffer from bias; specifically, they seem to heavily
underestimate $\phi$. $BC$ CIs do seem to correct this bias on average, but
the width of the CIs is very inconsistent. Our proposed CIs correct the bias
using $\widehat{\text{bias}} = \bar\theta_n^{(b)} -  \hat\theta_n$, which
seems to usually be negative, indicating a positive correction from interval
construction in Section~\ref{sec:bbci}.

@@

\jy{Keep the tex source tidy. Clean it up.}\mc{addressed}

Block bootstrap is a useful method for estimating parameters of a time series,
from simple parameters like the mean to more complicated temporal dependence
factors. We know theoretically that the block bootstrap procedure will cover
parameter of a time series at the nominal level given an infinitely large
sample, so the goal for this study was to find the smallest finite sample
length $n$ of a time series in order for the block bootstrap procedure to 
recover its associated parameters at an acceptable rate.

Again, this analysis relies on the assumption that there is a size $n$ large
enough for the method to work: that is, the method's performance improves as
$n$ increases. Out of the five types of intervals used in this study, this
assumption was found to hold true with respect to estimating $\phi$ only for
Standard, Student's $t$, and Proposed Method CIs, whereas Percentile, $BC$,
and $BC_a$ intervals exhibited coverage deterioration as $n$ increased. The
Percentile CI's coverage deterioration can be attributed to bias that is not
corrected as $n$ increases. Specifically, as $n$ increases, the width of the
CI decreases, but because the Percentile CI underestimates $\phi$, the
coverage decreases. The $BC$ CI seems to correct the bias, but the width of
the CI seems to be inconsistent. The acceleration factor of the $BC_a$ CI
seems to fail, as the width of the CI seems to be inconsistent.

When using Student's $t$ intervals and $\phi$ is unknown, the results of this
study suggest that an $n \geq 800$ may be necessary for common practice to
estimate simpler parameters like $\mu$ or $\sigma_x$. If $\phi$ is already
known, a lesser $n$ may be adequate to estimate these parameters. For lower
absolute values of $\phi$, a smaller $n$ is sufficient when estimating $\mu$
or $\sigma$. For negative values of $\phi$, coverage of $\mu$ is higher in
general. In fact, Student's $t$ seems to over-cover $\mu$ for $\phi = 0.4$
when using smaller sample sizes, meaning that if $\phi$ is known to have a
very negative value, another CI may be required. However, in real world
applications, $\phi$ is more commonly found to be positive, so Student's $t$
should still be used if $\phi$ is unknown. Lastly, to estimate $\phi$,
$n \geq 100$ using the Student's $t$ method may be sufficient. Further
investigation may be necessary to see if there are other interval corrections
that fix the coverage deterioration problem for $\phi$.

This study could be used as a guide for applied statistics courses for students
to generally understand how large of a sample size is sufficient for block
bootstrap to be used versus other inference methods. For undergraduate or
graduate students, block bootstrap is not typically a part of curriculum, but
the results of this study can easily be used to demonstrate when it is
practical to use this method. This information could also prove to be useful
for research using block bootstrap estimation of time series in domains such
as econometrics. Future studies could investigate if there are types of block
bootstrap interval construction such as $ABC$ or bootstrap-t intervals, or
alternatives to block bootstrap such as AR-Sieve bootstrap that would more
appropriately recover the parameters of a time series.
\eds{are there other methods that you know of?} \mc{mentioned an alternative
  to block bootstrap}
One could also investigate the $n$ needed to make inferences about other forms
of serially dependent data such as a moving average process.

>>>>

\paragraph{A Proposed CI}
\jy{Can we come up with a good name for the proposed method? Other people may cite this.}\mc{bias-centered? i'm not sure}
\jy{How about ``recentered percentile''? It has the same length as the
  percentile approach but is centered at the point estimate.}

We propose a CI that is centered at the point estimator and uses the variation
in the bootstrap estimates to construct the error bound. This interval requires
the computation of $\bar\theta_n^{(b)}$, the mean of all bootstrap point
estimates, and $\widehat{\text{bias}} = \bar\theta_n^{(b)} - \hat\theta_n$. A
$(1 - \alpha)100\%$ CI is
\[
(\hat\theta_{n, \alpha/2}^{(b)} - \widehat{\text{bias}}, \quad
\hat\theta_{n, 1 - \alpha/2}^{(b)} - \widehat{\text{bias}}).
\]
This CI is centered around $\hat\theta_n$ and can also be rewritten
as 
\[
(\hat\theta_n + \hat\theta_{n, \alpha/2}^{(b)} - \bar\theta_n^{(b)}, \quad
\hat\theta_n + \hat\theta_{n, 1 - \alpha/2}^{(b)} - \bar\theta_n^{(b)}).
\]
It is asymmetric, as different critical values are used to compute
the lower and upper bounds.

@@

\subsection{Design}
We designed a simulation study to compare the performance of the different
block bootstrap CI methods. In particular, we generated time series $X_t$ from
a 1st order auto-regressive (AR(1)) process:
\[
X_t = \phi X_{t-1} + \epsilon_t,
\]
where $\phi$ is an auto-regressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: $\{-0.4, -0.2, 0.0, 0.2, 0.4\}$. The
series $X_t$ has mean zero and variance
$\sigma_x^2 = (1 - \phi^2) \sigma_{\epsilon}^2$, so for each value of $\phi$,
we set $\sigma_{\epsilon}^2 = 1 / (1 - \phi^2)$ such that $\sigma_x^2 = 1$.

@@

\subsection{Results}

For estimating the mean parameter~$\mu$, Figure~\ref{fig:mu} suggests that all
methods eventually approach correct coverage of $\mu$ as sample size
increases. Student's $t$ CIs appear to need the smallest sample size
to achieve correct coverage, except for samples with strong negative
dependence, in which case, they actually over-cover $\mu$ for smaller sample
sizes. For instance, for a sample with $n = 100$ and $\phi = -0.4$, the lower
bound for a Students $t$ CIs coverage of $\mu$ is greater than 95\%, whereas
the coverage intervals for other methods contain 95\%. The standard,
Percentile, $BC$, and $BC_a$, and Proposed Method CIs require similar sample
sizes to to recover $\mu$ at the nominal level for all combinations of $n$ and
$\phi$. All methods seem to require a smaller sample to recover $\mu$ at the
nominal rate when dealing with negative dependence versus positive dependence.
For example, $BC$ CIs recovers $\mu$ for $n \geq 100$ when $\phi = -0.2$, but
they only recover $\mu$ for $n \geq 800$ when $\phi = 0.2$. In addition, as a
negative dependence gets stronger, holding everything else equal, coverage
increases, which lead to the Student $t$ CI's aforementioned over-coverage. As
a positive dependence gets stronger, holding everything else equal, coverage
decreases, and a larger sample is necessary to recover~$\mu$.

For estimating the standard deviation parameter~$\sigma_x$,
Figure~\ref{fig:sigma} suggests that every method can reach nominal coverage of
$\sigma_x$ if the sample is large enough, but for a given $n$ and $\phi$,
coverage of $\sigma_x$ will be lower than coverage of $\mu$ in general. Like
$\mu$, $\sigma_x$ can be covered by Student $t$ CIs with smaller sample sizes
when compared to other methods, and there is no over-coverage issue for
$\phi = -0.4$. Standard, percentile, $BC$, $BC_a$, and Proposed Method CIs
again have similar performance. All methods seem to have slightly higher
coverage of $\sigma_x$ when $\phi$ is negative versus when $\phi$ is positive.
Regardless of the sign, coverage of $\sigma_x$ gets worse as the strength of
the temporal dependence increases.
\jy{Compared to Figure 1, there is no over-coverage in this case. Say someting
  about this observation.}

@@

\jy{Use cross for red. Otherwise, they are not distinguishable when printed in b/w.}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/norm_phi_intervals}
  \caption{100 replicate Percentile, Bias-Corrected, and Proposed Method
    Bootstrap CIs for samples of $n = 800$ for an AR(1) process with
    $\phi = 0.4$. For each replicate, the lower and upper bounds of the CIs
    are displayed, as well as $\hat\theta_n$ (blue circle) and
    $\bar\theta_n^{(b)}$ (red cross). }
  \label{fig:npi}
\end{figure}

\jy{can you also do the same plot for sample size 1600? And then put the two
  plots in one figure?}

The outcomes of the $\phi$ estimation raise a natural question about the
lackluster performance of certain methodologies. To delve into this inquiry, a
set of 100 CIs was generated for each of the percentile,
$BC$, and Proposed techniques. Illustrated in Figure~\ref{fig:npi}, it becomes
evident that the percentile-based CIs exhibit a notable bias, predominantly
manifesting as a substantial underestimation of $\phi$. On the other hand, $BC$
CIs do display a tendency to rectify this bias on average, albeit accompanied by
a rather inconsistent width across the intervals. Our proposed CIs manage to
effectively mitigate the bias, employing $\widehat{\text{bias}} =
\bar\theta_n^{(b)} - \hat\theta_n$ as a corrective measure. Notably, this
correction often takes a negative form, indicative of a positive adjustment
arising from the interval construction detailed in Section~\ref{sec:bbci}.

To summarize, the performance of the CIs depends on the target parameter. When
estimating $\mu$ and $\sigma_x$, any CI will do, although Student's $t$ CIs
perform noticeably better than the others. However, when estimating $\phi$,
the choice of method is of utmost importance as to avoid coverage
deterioration. A smaller sample size is generally required to estimate a
parameter for a sample with a negative $\phi$ versus a positive $\phi$ of the
same magnitude. In order to know if coverage will increase as the strength of
the temporal dependence increases, one need to know what the parameter of
interest is, and in the case of $\mu$, the direction of the serial dependence.
\jy{The BC approach appears not correcting the bias correctly. The proposed
  approach simply does not correct point estimate and only uses the spreadness
  from the bootstrap to construct CIs.}

@@

Block bootstrap is a useful method for estimating parameters of a time series,
from simple parameters like the mean to more complicated temporal dependence
factors. We know theoretically that the block bootstrap procedure will cover
parameter of a time series at the nominal level given an infinitely large
sample, so the goal for this study was to find the smallest finite sample
length $n$ of a time series in order for the block bootstrap procedure to 
recover its associated parameters at an acceptable rate.
Our analysis relies on the assumption that there is a size $n$ large
enough for the method to work: that is, the method's performance improves as
$n$ increases. Out of the five types of intervals used in this study, this
assumption was found to hold true with respect to estimating $\phi$ only for
Standard, Student's $t$, and Proposed Method CIs, whereas Percentile, $BC$,
and $BC_a$ intervals exhibited coverage deterioration as $n$ increased. The
Percentile CI's coverage deterioration can be attributed to bias that is not
corrected as $n$ increases. Specifically, as $n$ increases, the width of the
CI decreases, but because the percentile CI underestimates $\phi$, the
coverage decreases. The $BC$ CI seems to correct the bias, but the width of
the CI seems to be inconsistent. The acceleration factor of the $BC_a$ CI
seems to fail, as the width of the CI seems to be inconsistent.

\jy{use a better topic sentence.}
When using Student's $t$ intervals and $\phi$ is unknown, the results of this
study suggest that an $n \geq 800$ may be necessary for common practice to
estimate simpler parameters like $\mu$ or $\sigma_x$. If $\phi$ is already
known, a lesser $n$ may be adequate to estimate these parameters. For lower
absolute values of $\phi$, a smaller $n$ is sufficient when estimating $\mu$
or $\sigma$. For negative values of $\phi$, coverage of $\mu$ is higher in
general. In fact, Student's $t$ seems to over-cover $\mu$ for $\phi = 0.4$
when using smaller sample sizes, meaning that if $\phi$ is known to have a
very negative value, another CI may be required. However, in real world
applications, $\phi$ is more commonly found to be positive, so Student's $t$
should still be used if $\phi$ is unknown. Lastly, to estimate $\phi$,
$n \geq 100$ using the Student's $t$ method may be sufficient. Further
investigation may be necessary to see if there are other interval corrections
that fix the coverage deterioration problem for $\phi$.

This study could be used as a guide for applied statistics courses for students
to generally understand how large of a sample size is sufficient for block
bootstrap to be used versus other inference methods. For undergraduate or
graduate students, block bootstrap is not typically a part of curriculum, but
the results of this study can easily be used to demonstrate when it is
practical to use this method. This information could also prove to be useful
for research using block bootstrap estimation of time series in domains such
as econometrics. Future studies could investigate if there are types of block
bootstrap interval construction such as $ABC$ or bootstrap-t intervals, or
alternatives to block bootstrap such as AR-Sieve bootstrap that would more
\jy{Cite references for AR-sieve}
appropriately recover the parameters of a time series.
\eds{are there other methods that you know of?} \mc{mentioned an alternative
  to block bootstrap}
One could also investigate the $n$ needed to make inferences about other forms
of serially dependent data such as a moving average process.
